### 2018-03-02 paper-tourrd3.Rmd
Small changes only

---
title: Connecting R with D3 for dynamic graphics, to explore multivariate data with tours
author:
  - name: Michael Kipp
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  mkipp271@gmail.com
  - name: Dianne Cook
    affiliation: Monash University
    address:
    - Department of Econometrics and Business Statistics
    email:  dicook@monash.edu 
  - name: Ursula Laa
    affiliation: Monash University
    address:
    - School of Physics and Astronomy
    email:  ursula.laa@monash.edu
bibliography: references.bib
link-citations: true
fig_caption: yes
abstract: >
  The tourr package in R has several algorithms and displays for showing multivariate data as a sequence of low-dimensional projections. It can display as a movie but has no capacity for interaction, such as stop/go, change tour type, drop/add variables. The tourrGui  package provides these sorts of controls, but the interface is programmed with the dated RGtk2 package. This work explores using custom messages to pass data from R to D3  for viewing, using the Shiny framework.  
preamble: >
  % Any extra latex you need in the preamble
output: rticles::rjournal_article
---

```{r cache=FALSE, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}
#packages:
library(bookdown)
library(knitcitations)
library(RefManageR)
knitr::opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, kfigr.link=TRUE, kfigr.prefix=TRUE, cache=TRUE, fig.env = TRUE, fig.cap=TRUE, fig.height=3)

options("citation_format" = "pandoc")

BibOptions(check.entries = FALSE, style = "markdown", bib.style = "alphabetic", cite.style = 'alphabetic')
```

```{r}
library(shiny)
library(tourr)
```

## Introduction

The tour algorithm [@gt_pp; @gt_pp_mc] is a way of systematically generating and displaying projections of high-dimensional spaces in order for the viewer to examine the multivariate distribution of data. It can do this either randomly, or by picking projections judged interesting according to some criterion or index function. The tourr package [@tourr] provides the computing and display in R [@R; @ihaka:1996] to make several types of tours: grand, guided, little and local. The projection dimension can be chosen between one and the number of variables in the data. The display, though, has no capacity for interaction. The viewer can watch the tour like a movie, but not pause it and restart, or change tour type, or number of variables.

These interactive controls were provided with the tourrGui package [@tourrGui], with was programmed with the RGtk2 package [@RGtk2]. This is not the toolkit of choice today, and has been superceded with primarily web-capable tools, like Shiny [@shiny]. To display dynamic graphics though, is not straight-forward. This paper explains how to use D3 [@D3] as the display engine in a Shiny GUI, using custom message passing between server and client.

## Creating a tour, with the tourr package

The \pkg{tourr} package [@tourr] is an R implementation of the tour algorithms discussed in @gt_pp_mc. It includes methods for geodesic interpolation and basis generation, as well as an implementation of the simulated annealing algorithm to optimise projection pursuit indices for the guided tour. The tour can be displayed directly in the R graphics device, for example:

```{r eval=FALSE, echo=TRUE}
# quartz() # to display on a Mac
# X11() # For windows
# The Rstudio graphics device cannot be used
animate_dist(flea[, 1:6], center = FALSE)
```

generating the animation in Figure \ref{tour}.

\begin{figure}[ht]
EMBED MOVIE HERE
\end{figure}

A tour path is a smooth sequence of projection matrices, $p\times d$, that when combined with a matrix of n data points, $n\times p$, and a rendering method, produces a steady stream of $d$-dimensional views of the data.  Each tour is initialised with the `new_tour()` method, which instantiates a tour object and takes as arguments the data $X$, the tour method, e.g. `guided_tour()`, and the starting basis. Once initialised, a new target plane is chosen, and a series of steps along a geodesic path from starting to target plane are generated by interpolation.

This requires a series of calls to the tour object producing the series of projections. The steps are discrete, of size given by $\omega/\Delta$, where $\omega$ denotes the angular velocity of the geodesic interpolation, and $\Delta$ is a parameter denoting frames per second, reflecting the rendering speed of the device in use. The $\Delta$ parameter can be thought of as the frames per second, while $\omega$ affects the speed at which the tour moves through the projection space. For our purposes, $\Delta$, `fps` in the code, is set at 25, while the $\omega$ can be adjusted by the user.

## Connecting the tour projections to D3 display using `sendCustomMessage`

D3.js (Data-Driven Documents) [@D3] is a JavaScript library for manipulating documents based on data. The advantages of D3 are similar to those provided by Shiny: namely, an industry standard with rich array of powerful, easy to use methods and widgets that can be displayed on a wide variety of devices, with a large user base. D3 works on data objects in the JavaScript Object Notation (JSON) format, which are then parsed and used to display customisable data visualisations.

The new implementation of the tour interface uses D3 to render each projection step returned by R, focusing on 2D projections as a test case. It does this by drawing and re-drawing a scatterplot with dots (or circles in D3 language) and providing SVG objects for the web browser to render.

There are two functions provided by the Shiny framework to transport data between R and JavaScript: `session$sendCustomMessage()` in R, and the corresponding `Shiny.addCustomMessageHandler()` in JavaScript. Whenever the former is executed in R, the latter function will execute a code block in JS. There are many examples of such functions being used to pass arbitrary data from an R app to a JS front-end, few examples exist of this basic functionality to update a D3 animation in real-time.

The data format expected by D3 is in JSON format, which combines two basic programming paradigms: a collection of name/value pairs, and an ordered list of values. R's preferred data formats include data frames, vectors and matrices. Every time a new projection has been calculated with the tour path, the resulting matrix needs to be converted to JSON and sent to D3. The code to send the D3 data looks like this:

```
session$sendCustomMessage(type = "data", message = toJSON(j))
```

This code is from the observe environment from the `server.R` file. It converts the matrix of projected data points to JSON format, and sends it to JavaScript with the id data. When parsed in D3 by its `data()` method, it is converted back into a logical 2D array where the columns are queried first, then the rows. If column names are included in the JSON, the column indices are strings; otherwise they are integers starting from 1.
All of the code required to render the scatterplots and legends, along with colours, is JavaScript code in the file `d3anim.js`. In particular, the data from R is handled with the following code:

```
Shiny.addCustomMessageHandler("data",
    function(message) {
        /* D3 scatterplot is drawn and re-drawn using the
            data sent from the server. */
}
```

Every time the message is sent (25 times per second), the code-block is run.

## Getting projections

The `observeEvent` Shiny method defines a code block to be run whenever some input value changes. The following code snippet restarts a tour using a random basis:

```
observeEvent(input$restart_random,
  {
    p <- length(input$variables)
    b <- matrix(runif(2*p), p, 2)
    rv$tour <- 
      new_tour(as.matrix(rv$d[input$variables]),
              choose_tour(input$type, 
              input$guidedIndex,
              c(rv$class[[1]]), 
              input$scagType), b)
})
```

The projections are calculated using the tour object in an `observe()` environment, which repeatedly runs the code until a reactive variable is changed, at which point it is "invalidated" and then re-started. The projections are calculated using the following code
block:

```
observe({
  tour <- rv$tour
  aps <- rv$aps
  step <- tour(aps / fps)
  if (!is.null(step)) {
    invalidateLater(1000 / fps)
    j <- center(rv$mat %*% step$proj)
    j <- cbind(j, class = rv$class)
    colnames(j) <- NULL
    session$sendCustomMessage(type = "data", 
      message = toJSON(j))
  }
  else {
    if (length(rv$mat[1, ]) < 3) {
    session$sendCustomMessage(type = "debug", 
      message = "Error: Need > 2 variables.")
    } else {
    session$sendCustomMessage(type = "debug", 
    message = "Guided tour finished: no better bases found.")
    }
  }
})
```

## Pros and cons

The D3 canvas makes for smooth drawing and re-drawing of the data projections. Adding a graphical user interface (GUI) around the display is straightforward with the Shiny package, e.g. control elements such as stop/go, increase/decrease speed, change tour type, add/remove variables from the mix.

The main disadvantage is that the speed is inconsistent, as server and client play tag to keep up with each other, and the display cannot handle many observations. Noticeable slow down was oberved with 2000 points, the main reason being the rendering time required for the large number of SVG circle elements. The situation can be improved when using a single HTML5 canvas element to draw the scatter points, significantly reducing the rendering time.

## Code

The code is available at [https://github.com/makipp/ TourrGuiD3], and the source material for this paper is available at [https://github.com/dicook/paper-tourrd3].

## Summary

The `sendCustomMessage` function provides a way to share a tour path with the D3 renderer, and embed it in a Shiny GUI providing controls such as stop/go, increase/decrease speed, change tour type, add/remove variables. However, the approach can result in inconsistent speed which appears as jerkiness in the motion, and it isn't suitable for large numbers of observations.

## References

<div id="refs"></div>